---
title: "p8105_hw3_xh2636"
author: "Xiaoyu Huang"
date: "2023-10-09"
output: github_document
---

# Problem 1
```{r}
library(p8105.datasets)
library(tidyverse)
library(ggplot2)
data("instacart")
```

```{r, include=FALSE}
data_dim <- dim(instacart)

# Print the number of observations and variables
num_obs <- nrow(instacart)
num_obs
num_var <- ncol(instacart)
num_var
```

The dataset has `r num_obs` observation and `r num_var` variables. The key variable in the dataset are their “ids”, “aisle”, and “department name”. Also, we can see the data set contain different product name.

```{r, include=FALSE}
# Find number of aisle
num_aisles <- length(unique(instacart$aisle))
num_aisles
```

```{r}
# Find number aisles are the most items ordered from and rank it
rank_aisle <- instacart %>%
  group_by(aisle) %>%
  summarize(total_ordered = n())%>%
  arrange(-total_ordered)
rank_aisle
```

The are `r num_aisles` aisles, and aisle have the most items ordered from is fresh vegetable.

```{r}
# limiting this to aisles with more than 10000 items ordered.
aisle_counts <- table(instacart$aisle)
popular_aisles <- names(aisle_counts[aisle_counts > 10000])
```

```{r}
# Make a plot that shows the number of items ordered in each aisle
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  arrange(desc(n)) %>%
  mutate(aisle = fct_reorder(aisle, n)) -> filtered_data

aisle_order_plot <- ggplot(filtered_data, aes(x = aisle, y = n)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  labs(x = "Aisle", y = "Number of Items Ordered") +
  theme_minimal()

print(aisle_order_plot)
```

We can see the arrangement of these data from high to low, and we can see that there are 39 categories with more than 10,000 quantities, and fresh vegetable has the largest number. Butter accounts for the smallest proportion here.

```{r, message=FALSE}
# Make a table showing the three most popular items
three_mpi <- instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits"))

# Group aisle and product name
aisle_product_counts <- three_mpi %>%
  group_by(aisle, product_name) %>%
  summarise(order_count = n())%>%
  arrange(aisle, desc(order_count))

# Rank the items within each aisle by order count
aisle_product_counts <- aisle_product_counts %>%
  group_by(aisle) %>%
  mutate(rank = rank(-order_count))
top_items <- aisle_product_counts %>%
  filter(rank <= 3)
top_items <- top_items %>%
  select(aisle, rank, product_name, order_count)

print(top_items)
```

We selected the top three comprehensive rankings from these three categories, so that we can read this table more intuitively. And we can see that the number of vegetable orders is the highest, far exceeding the number of dog food care.

```{r, message=FALSE, warning=FALSE}
hour_day <- instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream"))

# Extract day of the week and hour of the day from the "order_hour_of_day" column
hour_day <- hour_day %>%
  mutate(day_of_week = weekdays(as.Date(order_dow, origin = "1990-01-01")),
         hour_of_day = order_hour_of_day)

# Calculate the mean hour of the day for each product on each day of the week
myday <- hour_day %>%
  group_by(product_name, day_of_week) %>%
  summarise(mean_hour = mean(hour_of_day, na.rm = TRUE)) %>%
  pivot_wider(names_from = day_of_week, values_from = mean_hour)

# Specify the order of days of the week
day_order <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
myday <- myday %>%
  select(product_name, day_order)
print(myday)
```

From the data we compiled, we can see that for coffee ice cream, Tuesday has the largest average value and Friday has the smallest average value. On the other hand, pink lady apple has the largest average on Tuesday and the smallest on Sunday.

#Problem 2


